Author: Jason Bryer, Ph.D. (jason@bryer.org)  
Website: [bryer.org](http://jason.bryer.org)

### Running the App

The app is deployed to RStudio's [shinyapps.io](https://jbryer.shinyapps.io/BayesBilliards/) service at [jbryer.shinyapps.io/BayesBilliards](https://jbryer.shinyapps.io/BayesBilliards).

The Shiny App is included in the [`DATA606`](https://github.com/jbryer/DATA606) package on Github and can be run, once installed, using the `DATA606::shiny_demo('BayesBilliards')` function.

Or, run the app directly from Github using the `shiny::runGitHub('DATA606', 'jbryer', subdir='inst/shiny/BayesBilliards')` function.

### Problem Statement

Consider a pool table of length one. An 8-ball is thrown such that the likelihood of its stopping point is uniform across the entire table (i.e. the table is perfectly level). The location of the 8-ball is recorded, but not known to the observer. Subsequent balls are thrown one at a time and all that is reported is whether the ball stopped to the left or right of the 8-ball. Given only this information, what is the position of the 8-ball? How does the estimate change as more balls are thrown and recorded?

### Usage

* **Confidence Range** - The width of the confidence interval around the mean.
* **Show Pool Table** - Show a figure with the 8-ball along with the location of subsequent ball throws.
* **Distribution Plot Type** - For the proir and posterior distribution plots use density or histogram plot types.
* **Next Ball** - Throw another ball and record whether it fell to the left or right of the 8-ball.
* **Start Over** - Throw a new 8-ball and start the process over.
* **Adjust number of samples per iteration** - This will provide a slider input to adjust the number of the simulated random samples for each iteration.
* **Number of Samples** - The length of the prior distributions. For the first iteration, *k* random numbers are drawn from the uniform distribution. For subsequent iterations, *k* random numbers are drawn the posterior distribution of the previous iteration, with replacement so to have a prior of length *k* with the same shape for the current iteration.

### Strategy

For the initial iteration, we have one observation of left or right. We assume a uniform *prior* distribution of length *k*. For each *k*, we sample `unif(0,1)` and record whether that value is less than (i.e. left) or greater than (i.e. right) of *k*. The values from *k* that match the original observation remain and constitutes our *posterior* distribution.

For subsequent iterations, we simply use our *posterior* distribution from the prior iteration as our *prior* distribution for the current sampling.

In R, we start with a vector `prior` drawn from `runif`.

```{r, message=FALSE, warning=FALSE, error=FALSE, fig.width=6, fig.height=4}
k <- 5000
prior <- runif(k, min=0, max=1)
hist(prior)
```

Assume the first ball falls to the left of the 8-ball. We randomly draw *k* values from a uniform distribution and keep the values from the prior distribution where the random value is greater. This is the posterior distribution.

```{r, message=FALSE, warning=FALSE, error=FALSE, fig.width=6, fig.height=4}
posterior <- prior[prior < runif(k, min=0, max=1)]
hist(posterior)
```

Now, consider we throw another ball that lands to the right. The prior distribution for this iteration is the posterior from the previous iteration. In order to have *k* elements in the prior distribution, we sample *k* from the posterior with replacement.

```{r, message=FALSE, warning=FALSE, error=FALSE}
prior <- sample(posterior, k, replace=TRUE)
```

Getting the posterior distribution is the same the initial iteration.

```{r, message=FALSE, warning=FALSE, error=FALSE, fig.width=6, fig.height=4}
posterior <- prior[prior > runif(k, min=0, max=1)]
hist(posterior)
```

With two observations, there is a 50% chance the 8-ball is between `r quantile(posterior, c(.25, .5))[1]` and `r quantile(posterior, c(.25, .5))[2]`.

```{r, message=FALSE, warning=FALSE, error=FALSE, fig.width=6, fig.height=4}
mean(posterior)
quantile(posterior, c(.25, .5))
```


### References

Bayes, T. (1763), An essay towards solving a problem in the Doctrine of Chances. *Philosophical Transactions of the Royal Society of London, 53*. Retrieved from http://www.stat.ucla.edu/history/essay.pdf

Downey, A. (2015). Bayesian Billiards. Retrieved from http://allendowney.blogspot.com/2015/06/bayesian-billiards.html

Eddy, S.R. (2004). What is Bayesian statistics? *Nature Biotechnology 22*. Retrieved from http://www.nature.com/nbt/journal/v22/n9/full/nbt0904-1177.html

McGrayne, S.B. (2011). *The Theory That Would Not Die: How Bayes' Rule Cracked the Enigma Code, Hunted Down Russian Submarines, and Emerged Triumphant from Two Centuries of Controversy.* Yale University Press.


